\chapter{Introduction}
\todo[inline]{Jos: Add architectural diagrams and make the attacker models clearer}
\todo[inline]{Jos: Can be given more detail more detail with some subsections}
\todo[inline]{Jos: One typical example}
\todo[inline]{Rewrite}
\todo[inline]{Luko: paragraaf twee van de introductie voelt voor mijn een beetje alsof hij uit het niks valt}
\todo[inline]{Geert: Add something about CIS}
\todo[inline]{Geert: Footnotes after punctuation}
\todo[inline]{Geert: Add page number references besides section references}
\todo[inline]{Geert: citations at the end of sentence after the punctuation}
\todo[inline]{Geert: No monospaced titles}
\todo[inline]{Think about monospace usage}
\todo[inline]{Geert: Sort citations}
Containerization software allows developers to package software in small, easily reproducible packages and run software in lightweight, isolated environments. Because of the ease of use and isolation it brings, containerization has become popular for development and deployment of software. This is why many companies have started using it.

Secura, a company specializing in digital security, performs security assessments for clients. In these assessments, Secura evaluates the private and public networks of their clients. Secura would like to improve those assessments by also looking into containerization software that is used by a selection of their clients. This thesis focuses on the techniques needed to perform an effective assessment of systems that use Docker.

\medskip

We will first look at the necessary concepts (\autoref{chapter:notation}) and background information (\autoref{chapter:background}) on containerization software and Docker, the de facto industry standard for containerization software and focus of this thesis. We will then go into more detail about the attack surface (\autoref{chapter:attack-surface-models}), misconfigurations (\autoref{section:misconfigurations}) and specific security related bugs (\autoref{section:bugs}). We will discuss how these can be identified during a penetration test (\autoref{chapter:pentesting}). Most importantly, we will combine all information into a checklist of questions that penetration testers can use to test the security of systems that use Docker (\autoref{chapter:checklist}).

Finally, we will look at out of scope but interesting ideas to extend this research (\autoref{chapter:futurework}), other research about security and Docker (\autoref{chapter:relatedwork}) and the takeaways of this thesis from both an offensive and a defensive perspective (\autoref{chapter:conclusions}).

\medskip

We will focus on Linux, because Docker is developed for Linux (although non-Linux Docker versions do exist\footnote{Docker on non-Linux systems runs inside a Linux virtual machine.}). Throughout this thesis we will look at practical examples, so a good understanding of Linux is helpful.
