\subsection{Host Firewall Bypass}\label{subsection:iptables}
The Linux kernel has a built-in firewall. This firewall consists of multiple chains of rules which are stored in tables. Each table has a different purpose. For example, there is a \lstinline{nat} table for address translation and a \lstinline{filter} table for traffic filtering (which is the default).
Each table has chains of ordered rules which also have a different purpose. For example, there are the \lstinline{OUTPUT} and \lstinline{INPUT} chains in the \lstinline{filter} table that are meant for all outgoing and incoming traffic, respectively.
It is possible to configure these rules using a program called \lstinline{iptables}. All Linux based firewalls (e.g. \lstinline{ufw}) use \lstinline{iptables} as their backend.

\hfill

When the Docker Daemon is started, it sets up its own chains and rules to create isolated networks. The way it sets up its rules completely bypasses other in the firewall (because they are setup before the other rules) and by default the rules are quite permissive. This is by design, because the network stack of the host and the container are separate, including the firewall rules. It is, however, a bit counterintuitive, because one would assume that if a firewall rule is set on the host, it would apply to everything running on that host including containers (and virtual machines).

\hfill

We will look at the following simple example of bypassing a firewall rule with Docker.

\begin{lstlisting}
(host)# iptables -A OUTPUT -p tcp --dport 80 -j DROP
(host)# iptables -A FORWARD -p tcp --dport 80 -j DROP
(host)$ curl http://httpbin.org/get
curl: (7) Failed to connect to httpbin.org port 80: Connection timed out
(host)$ docker run -it --rm ubuntu /bin/bash
(cont)# apt update
(cont)# apt install curl
(cont)# curl http://httpbin.org/get
{
  "args": {},
  "headers": {
    "Accept": "*/*",
    "Host": "httpbin.org",
    "User-Agent": "curl/7.58.0"
  },
  "origin": "<redacted>",
  "url": "https://httpbin.org/get"
}
\end{lstlisting}

We first setup rules to drop all outgoing (including forwarded) \lstinline{HTTP} (not \lstinline{HTTPS}) traffic. We drop traffic going to port \lstinline{80} (the default \lstinline{HTTP} port) and try to request a \lstinline{HTTP} page on the host. As expected, it does not work. If we then try to make the exact same request in a container, it works.

\hfill

The Docker CIS Benchmark does not cover this problem. It, however, does have a guideline that ensures this problem exists. Guideline 2.3 (Ensure Docker is allowed to make changes to iptables) recommends that the Docker Daemon is allowed to change the firewall rules. This is a good recommendation, because it removes the need to configure all network configuration for Docker yourself. But that also isolates the firewall rules of the host from the containers.
